---
title: "MD Ch. 3-4 Homework"
author: "Samuel Snelson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Chapter 3

```{r}
library(tidyverse)
theme_set(theme_light(base_family = "Avenir"))

```


Importing Mario kart world record dataset. 

```{r}

mario_kart <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/world_records.csv")

glimpse(mario_kart)
```

## Question 1

```{r}
three_laps <- mario_kart %>% 
  filter(type == "Three Lap")
```

Datasets with and without the Rainbow Road track: 

```{r}
three_without_rainbow <- three_laps %>% 
  filter(track != "Rainbow Road")

three_only_rainbow <- three_laps %>% 
  filter(track == "Rainbow Road")

```

## Question 2

Summary statistics for races at Rainbow Road and all other tracks 

```{r}

rainbow_summary <- three_only_rainbow %>% 
  summarize(mean = mean(time), std_dev = sd(time))

rainbow_summary

all_other_summary <- three_without_rainbow %>% 
  summarize(mean = mean(time), std_dev = sd(time))

all_other_summary

```

**The average time to complete a race is considerably longer on Rainbow Road that on every other track. That the standard deviation for Rainbow Road is larger than that of all others (grouped together) suggests that the difference in means isn't due to variability in completion times on other tracks but that Rainbow Road is just longer and/or harder to finish.**

## Question 3

Grouping records by track

```{r}
#Grouped by track, producing summary stats on the number of records for each track
summary_by_track <- three_laps %>% 
  group_by(track) %>% 
  summarize(mean = mean(time), std_dev = sd(time), n_records = n()) %>% 
  arrange(desc(n_records))

summary_by_track
```

**While perhaps an odd metric, Toad's Turnpike has had the most records set on it with 124 (in whatever period of time and scope of record keeping it refers to).**

## Question 4

Players who have multiple records on the same track(s)

```{r}
# Grouped by player and track, finding the number of records each player has on each track
player_records <- three_laps %>% 
  group_by(player, track) %>% 
  summarize(n_records = n()) %>% 
  arrange(desc(n_records))

player_records

```

**It appears that Penev holds the most records established with 26 on Choco Mountain - quite a way to spend one's time.** 

## Question 5

Average time for each track

```{r}
# Grouped by track, finding average time it takes to finish race on each track
average_times <- three_laps %>% 
  group_by(track) %>% 
  summarize(avg_time = mean(time)) %>% 
  arrange(desc(avg_time))

average_times
```

**Surprise, Surprise, the track with the highest average time is Rainbow Road with 276 (I presume) seconds.** 

Best time for each track

```{r}
# grouped by track, best (minimum) time it takes to complete race
best_times <- three_laps %>% 
  group_by(track) %>% 
  summarize(best_time = min(time)) %>% 
  arrange(best_time) %>% 
  select(track, best_time)

best_times
```

**The fasted record is 14.6 (presumably) seconds set on Wario Stadium. We could see who set the record by simply selecting for player in addition to the track and time.**

## Question 6

Creating a new variable for records lasting longer or shorter than 100 days before being usurped! 

```{r}
three_laps <- three_laps %>% 
  mutate(lasting_record = if_else(record_duration > 100, 1, 0))

three_laps
```

How many long lasting records does each player have and which player has the longest lasting record? 

```{r}
three_laps <- three_laps %>% 
  group_by(player) %>% 
  summarize(n_records = sum(lasting_record)) %>% 
  arrange(desc(n_records))
```

**See table (or tibble ;) ) for each players number of long lasting records. NR holds the most long lasting records with 81.** 

## Question 7

Merging where players are from onto record data

```{r}
drivers <- read_csv("/Users/samuelsnelson/PhD Materials /Year One /Fall 2022/Statistics/soc_722/Data /drivers.csv")

head(drivers)
head(three_laps)

three_laps_country <- left_join(three_laps, drivers, by = "player")
```

```{r}
records_by_country <- three_laps_country %>% 
  group_by(nation) %>% 
  summarize(n_records = n()) %>% 
  arrange(n_records) %>%
  drop_na() %>% 
  ggplot(aes(reorder(nation, n_records),
             n_records)) + 
  geom_bar(stat = "identity", color = "steelblue", fill = "steelblue") + 
  labs(title = "Three-Lap Mario Kart Race World Records Across Country", x = "Country", y = "Number of Records Held", caption = "Note: Data represents records held for longer than 100 days.") +
  scale_y_continuous(breaks = seq(0, 650, by = 50))

records_by_country
# I don't entirely like how this looks. The vertical scaling is okay. (1) The 0 value above the x axis (2) and the plot background would be better simpler
```

# Chapter 4

We've already got tidyverse loaded so let's get crack-a-lackin. 

## Question 1

```{r}
nfl_salary <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/nfl_salaries.csv")

head(nfl_salary)
```

## Question 2

Pivoting data into tidy format

```{r}
nfl_salary_tidy <- nfl_salary %>% 
  pivot_longer(names_to = "position", # column names to position
               values_to = "salary", 
               cols = -year) # want to keep year the same 

nfl_salary_tidy
```

## Question 3

```{r}
nfl_salary_tidy %>% 
  filter(position == "Quarterback") %>%
  ggplot(aes(salary)) + 
  geom_histogram(nrow = 2) + # Not sure why this isn't adjusting to 2 rows? 
  scale_x_continuous(labels = scales::label_number(suffix = "m", scale = 1e-6)) +
  facet_wrap(~ year) + 
  labs(title = "Distributions of NFL Quarterbacks' Income, 2011-2018", x = "Salary in Millions", y = "Number of Players")

# Trying to adjust scales of things but its a bit wonky. I need to do a piecewise sort of things where values less than 1 million are labeled more granularly and values over 1m are in 5 or 10 million

```

**The salary distribution for NFL quarterbacks, while always right skewed, is inching slightly towards less payers earning less and more players earning more.** 

## Question 4

Average salary for each position each year

```{r}
average_sal_years <- nfl_salary_tidy %>% 
  group_by(position, year) %>% 
  summarize(avg_sal = mean(salary, na.rm = TRUE)) %>% 
  arrange(desc(avg_sal))

# Checking missing values
#library(visdat)
#vis_dat(nfl_salary_tidy)
#vis_miss(nfl_salary_tidy)

```

## Question 5

```{r}
#mean(unlist(drop_na(nfl_salary_tidy[(nfl_salary_tidy[,"position"] == "Quarrback") & (nfl_salary_tidy[, "year"] == 2018), "salary"])))
```

Line graph of each positions average salary over time

```{r}
average_sal_years %>% 
  ggplot(aes(year, avg_sal, color = position)) +
  scale_y_continuous(labels = scales::label_number(suffix = "m", scale = 1e-6)) +
  labs(title = "Average Salary by NFL Position, 2011-2018", y = "Average Salary in Millions") +
  geom_smooth(size = 1, se = FALSE) 
```

**We see that both offensive and defensive linemen have dominates earnings in this period, now earning upwards of $7 million on average. Notably, it appears that the average salary of running backs have actually decreased from 2012 to 2015 and their 2012 height in 2018**
